---
title: "Fun with Sortable.js in R"
author: "Kenton Russell"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r echo = F}
### get knitr just the way we like it

knitr::opts_chunk$set(
  message = FALSE
  , warning = FALSE
  , error = FALSE
  , tidy = FALSE
  , cache = FALSE
  #, results = 'asis'
)
```

With `sortableR` [`htmlwidget`](http://htmlwidgets.org) we can amaze our non-R friends and thrill our colleagues with powerful, dependency-free interactivity from [`Sortable.js`](http://rubaxa.github.io/Sortable/) in the browser, RStudio Viewer, or RStudio Shiny.  I encourage you to **think creatively and imaginatively** with `sortableR` as there are all sorts of potential uses for this `htmlwidget`.  Please do not limit yourself to the simple examples within this vignette.

### Getting Started
`sortableR` has not achieved CRAN status yet, so for now we will install with `devtools::install_github` as shown in the code below.

```{r eval = F}
devtools::install_github("timelyportfolio/sortableR")
```

### Super Simple Example
Using `sortableR` in markdown is a little tricky since markdown does not provide an easy way to provide an `id` that we'll need.  We can overcome this by using bare `HTML` or using `htmltools::tags`.  Let's make a simple `ul` list.  Note, however, that `sortableR` works with nearly any `HTML` element, such as `div`.

```
<ul id = "example01">
  <li><h4>drag me</h4></li>
  <li><h4>sort me</h4></li>
  <li><h4>any way you like</h4></li>
</ul>
```

<ul id = "example01">
  <li><h4>drag me</h4></li>
  <li><h4>sort me</h4></li>
  <li><h4>touch (yes touch) me</h4></li>
  <li><h4>any way you like</h4></li>
</ul>

```{r}
library(sortableR)
sortableR( "example01" )
```

Now let's use the `htmltools::tags` approach to do exactly the same thing.

```{r}
library(htmltools)
tagList(
  tags$ul(id = "example02"
    ,tags$li(tags$h4("drag me"))
    ,tags$li(tags$h4("sort me"))
    ,tags$li(tags$h4("touch (yes touch) me"))
    ,tags$li(tags$h4("any way you like"))
  )
  ,sortableR( "example02" )
)
```

### Little Harder but Better Example

If you were wondering if `sortableR` works with images, of course it does.  We'll use `base64::img` to embed our plots on the page.

```{r}
library(base64)

# use example from ?base64::img
pngfile <- tempfile()
png( pngfile, width = 400, height = 300 )
plot( 1:100, rnorm(100), pch = 21, bg = "red", cex = 2 )
title(main="Best Scatter Chart Ever (Moves Like Jagger)")
msg<-capture.output(dev.off())


# make another plot for demo purposes
pngfile2 <- tempfile()
png( pngfile2, width = 400, height = 300 )
barplot( 1:9, col = blues9 )
title(main="Best Bar Chart Ever (I Like the Way You Move)")
msg<-capture.output(dev.off())

tagList(
  tags$div(id = "example03"
    ,HTML(img( pngfile ))
    ,HTML(img( pngfile2 ))
  )
  ,sortableR("example03")
)
```

```{r echo = F}
unlink(pngfile)
unlink(pngfile2)
```


### Power of Groups
Looking at the [`Sortable.js`](http://rubaxa.github.io/Sortable) excites me about the potential to use `sortableR` as an important UI element in both a Shiny and non-Shiny context.  We could potentially demo a plot builder with something like this example.  You'll notice that it doesn't really do anything, but I hope the intent and direction is clear.

```{r}
library(shiny)
fluidPage(
  fluidRow(class="panel panel-success"
    ,tags$div(class="panel-heading"
      ,tags$h3("Magic Can Happen Here")
    )
    ,fluidRow(class = "panel-body"
      ,column(width = 4
        ,tags$div(class="panel panel-default"
          ,tags$div(class="panel-heading","Variables")
          ,tags$div(class="panel-body"
            ,id = "sort1"
            ,lapply(
              colnames(mtcars)
              ,function(co){
                tag("p"
                  ,list(
                    class = class(mtcars[,co])
                    ,tags$span( class = "glyphicon glyphicon-move" )
                    ,tags$strong( co  )
                  )
                )
              }
            )      
          )
        )       
      )
      ,column(width =4
        ,tags$div(class="panel panel-default"
          ,tags$div(class="panel-heading"
            ,tags$span( class = "glyphicon glyphicon-stats" )
            ,"Analyze as x (drag here)"
          )
          ,tags$div(class="panel-body"
            ,id = "sort2"
          )
        )
      )
      ,column(width =4
        ,tags$div(class="panel panel-default"
          ,tags$div(class="panel-heading"
            ,tags$span( class = "glyphicon glyphicon-stats" )
            ,"Analyze as y (drag here)"
          )
          ,tags$div(class="panel-body"
            ,id = "sort3"
          )
        )            
      )
    )
  )
  ,sortableR( "sort1", options = list( group = "sortGroup1" ) )
  ,sortableR( "sort2", options = list( group = "sortGroup1" ) )
  ,sortableR( "sort3", options = list( group = "sortGroup1" ) )
)
```

### Code For a Shiny Example
I learned late last night that `sortableR` will allow movable tabs inside a Shiny (and also not Shiny) app.  By adding just one line of code and an `id` to this RStudio [Tabset example](https://github.com/rstudio/shiny-examples/tree/master/006-tabsets), we get rearrangeable tabs.  I really hope you'll copy and paste to see it for yourself, or you can `runGist("2dbe45f77b65e28acab9")`.

```{r eval=F}
# all credit for code goes to RStudio
# https://github.com/rstudio/shiny-examples/tree/master/006-tabsets
library(sortableR)
library(shiny)

ui = # Define UI for random distribution application 
shinyUI(fluidPage(
    
  # Application title
  titlePanel("Tabsets"),
  
  # Sidebar with controls to select the random distribution type
  # and number of observations to generate. Note the use of the
  # br() element to introduce extra vertical spacing
  sidebarLayout(
    sidebarPanel(
      radioButtons("dist", "Distribution type:",
                   c("Normal" = "norm",
                     "Uniform" = "unif",
                     "Log-normal" = "lnorm",
                     "Exponential" = "exp")),
      br(),
      
      sliderInput("n", 
                  "Number of observations:", 
                   value = 500,
                   min = 1, 
                   max = 1000)
    ),
    
    # Show a tabset that includes a plot, summary, and table view
    # of the generated distribution
    mainPanel(
      tabsetPanel(type = "tabs", id = "sortTab",
        tabPanel("Plot", plotOutput("plot")), 
        tabPanel("Summary", verbatimTextOutput("summary")), 
        tabPanel("Table", tableOutput("table"))
      )
    )
  )
  ,sortableR("sortTab")
))

server = function(input, output) {
  
  # Reactive expression to generate the requested distribution.
  # This is called whenever the inputs change. The output
  # functions defined below then all use the value computed from
  # this expression
  data <- reactive({
    dist <- switch(input$dist,
                   norm = rnorm,
                   unif = runif,
                   lnorm = rlnorm,
                   exp = rexp,
                   rnorm)
    
    dist(input$n)
  })
  
  # Generate a plot of the data. Also uses the inputs to build
  # the plot label. Note that the dependencies on both the inputs
  # and the data reactive expression are both tracked, and
  # all expressions are called in the sequence implied by the
  # dependency graph
  output$plot <- renderPlot({
    dist <- input$dist
    n <- input$n
    
    hist(data(), 
         main=paste('r', dist, '(', n, ')', sep=''))
  })
  
  # Generate a summary of the data
  output$summary <- renderPrint({
    summary(data())
  })
  
  # Generate an HTML table view of the data
  output$table <- renderTable({
    data.frame(x=data())
  })
  
}

shinyApp( ui, server )
```


### Conclusion
**Think big with this little `htmlwidget`**.

### Thanks
With powerful tools from generous people, amazing things can be easily accomplished.  Thanks to:

- [R](http://r-project.org)
- [Ramnath Vaidyanathan](https://github.com/ramnathv) and [RStudio](http://rstudio.com) for [`htmlwidgets`](http://htmlwidgets.org)
- [\@RubaXa](https://github.com/RubaXa) for [`Sortable`](https://rubaxa.github.io/Sortable)
