---
title: "Fun with Sortable.js in R"
author: "Kenton Russell"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r echo = F}
### get knitr just the way we like it

knitr::opts_chunk$set(
  message = FALSE
  , warning = FALSE
  , error = FALSE
  , tidy = FALSE
  , cache = FALSE
  #, results = 'asis'
)
```

With `sortableR` [`htmlwidget`](http://htmlwidgets.org) we can amaze our non-R friends and thrill our colleagues with powerful, dependency-free interactivity from [`Sortable.js`](http://rubaxa.github.io/Sortable/) in the browser, RStudio Viewer, or RStudio Shiny.  I encourage you to **think creatively and imaginatively** with `sortableR` as there are all sorts of potential uses for this `htmlwidget`.  Please do not limit yourself to the simple examples within this vignette.

### Getting Started
`sortableR` has not achieved CRAN status yet, so for now we will install with `devtools::install_github` as shown in the code below.

```{r eval = F}
devtools::install_github("timelyportfolio/sortableR")
```

### Super Simple Example
Using `sortableR` in markdown is a little tricky since markdown does not provide an easy way to provide an `id` that we'll need.  We can overcome this by using bare `HTML` or using `htmltools::tags`.  Let's make a simple `ul` list.  Note, however, that `sortableR` works with nearly any `HTML` element, such as `div`.

```
<ul id = "example01">
  <li><h4>drag me</h4></li>
  <li><h4>sort me</h4></li>
  <li><h4>any way you like</h4></li>
</ul>
```

<ul id = "example01">
  <li><h4>drag me</h4></li>
  <li><h4>sort me</h4></li>
  <li><h4>touch (yes touch) me</h4></li>
  <li><h4>any way you like</h4></li>
</ul>

```{r}
library(sortableR)
sortableR( "example01" )
```

Now let's use the `htmltools::tags` approach to do exactly the same thing.

```{r}
library(htmltools)
tagList(
  tags$ul(id = "example02"
    ,tags$li(tags$h4("drag me"))
    ,tags$li(tags$h4("sort me"))
    ,tags$li(tags$h4("touch (yes touch) me"))
    ,tags$li(tags$h4("any way you like"))
  )
  ,sortableR( "example02" )
)
```

### Little Harder but Better Example

If you were wondering if `sortableR` works with images, of course it does.  We'll use `base64::img` to embed our plots on the page.

```{r}
library(base64)

# use example from ?base64::img
pngfile <- tempfile()
png( pngfile, width = 400, height = 300 )
plot( 1:100, rnorm(100), pch = 21, bg = "red", cex = 2 )
title(main="Best Scatter Chart Ever (Moves Like Jagger)")
msg<-capture.output(dev.off())


# make another plot for demo purposes
pngfile2 <- tempfile()
png( pngfile2, width = 400, height = 300 )
barplot( 1:9, col = blues9 )
title(main="Best Bar Chart Ever (I Like the Way You Move)")
msg<-capture.output(dev.off())

tagList(
  tags$div(id = "example03"
    ,HTML(img( pngfile ))
    ,HTML(img( pngfile2 ))
  )
  ,sortableR("example03")
)
```

```{r echo = F}
unlink(pngfile)
unlink(pngfile2)
```
