---
title: "Using custom styles with CSS"
output: rmarkdown::html_vignette
description: >
  Sometimes you may want to control the appearance of your rank_list elements, and override the default CSS classes.
vignette: >
  %\VignetteIndexEntry{Using custom styles with CSS}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(sortable)
```

## Using custom styles

Sometimes you want to take control of the styling of your app, e.g. changing colours or sizes.

The functions `rank_list()` and `bucket_list()` allows you to set a custom CSS class, and you can then define your own CSS rules that modifies your shiny app.

The default CSS style for a rank list is `default-sortable`, but you can specify your own class by modifying the `class` argument.

For example, to add `custom-sortable` to the CSS class, use:

```r
rank_list(
  class = c("default-sortable", "custom-sortable"), # add custom style
  ...
)
```

Then you can use any of the standard techniques to [Style your apps with CSS](https://shiny.rstudio.com/articles/css.html) to modify your styling.

For example, to modify the `custom-sortable` class using a `shiny` tag, use:

```r
tags$style(
  HTML("
    .rank-list-container.custom-sortable {
      background-color: #8A8;
    }
    .custom-sortable .rank-list-item {
      background-color: #BDB;
    }
  ")
)
```

This app contains a fully worked example that turns the background colours of the `rank_list` into shades of green:


```{r, echo=FALSE}
library(htmltools)
tags$div(
  class = "shiny-app-frame",
  tags$iframe(
    src = "https://andrie-de-vries.shinyapps.io/sortable_custom_css_app/",
    width = "100%",
    height = 450
  )
)
```



## Source code

```{r echo=FALSE, cache=FALSE}
knitr::read_chunk(
  system.file("shiny-examples/custom_css/app.R", package = "sortable")
)
```

```{r custom-css-app, eval=FALSE}
```

